---
title: "Image Distributor"
output: null
---

This script copies annotated images from the main data folder (cpics_img) into corresponding folders in another directoty: diatoms, acantharea, copedos, etc. Use Section 1 for this function.

For copying annotated images into a single folder (not aggregated by class), use Section 2.

Created: June 18th 2024
By: Enrique Montes

# Load taxa lists and metadata table
```{r}
library(tidyverse)
library(lubridate)

# specify the directory where the files are located
dir_path <- "~/Library/CloudStorage/GoogleDrive-enriquemontes01@gmail.com/My Drive/GDrive/OCED_AOML/WS_cruises/plankton_imaging/CPICS/TS.Master_selection"

# obtain a list of file names in the directory
file_names <- list.files(path = dir_path, pattern = ".txt", full.names = TRUE)

# loop over each file and import the tables (use this for DATES)
for (file in file_names) {
  table_name <- gsub(".txt", "", basename(file)) # get the name of the table from the file name
  assign(table_name, read.table(file = file, header = FALSE, sep = "\t") %>%
           mutate(date = as.POSIXct(substr(V1, start = 24, stop = 38), format="%Y%m%d_%H%M%S", tz="UTC")))
}

# List of folder names with files lists
data_frame_names <- ls(pattern = "^class\\.")  # Get all data frames starting with "class."
```

# Set path to and create main directories (base directory where all images are and destination directory)
```{r}
path_to_files <- "~/Desktop/cpics_img/" # Base directory for all images
selected_base_dir <- "~/Desktop/selected_per_class/"  # Base directory for selected files in separate folders
selected_destination_dir_all <- "~/Desktop/selected2/"  # Base directory for selected files in common folder

# Create the base directory for selected files if it doesn't exist
if (!dir.exists(selected_base_dir)) {
  dir.create(selected_base_dir)
}

if (!dir.exists(selected_destination_dir_all)) {
  dir.create(selected_destination_dir_all)
}

# List all image files in all folders within the main directory
image_files <- list.files(path = path_to_files, pattern = "\\.(png|jpg|jpeg|tif|tiff)$", recursive = TRUE, full.names = TRUE)

# Count the number of image files
num_image_files <- length(image_files)

# Print the count of image files
cat("Number of image files in base directory:", num_image_files, "\n")
```

# Loop through each data frame (lists of annotated images) and copy them all in a common folder
```{r}
# Initialize an empty list to keep track of missing files
missing_files <- list()

# Loop through each data frame
for (df_name in data_frame_names) {
  # Get the data frame using the name
  current_df <- get(df_name)
  
  # Rename filename column if needed
  current_df <- current_df %>% rename(img_file_name = V1)
  
  # Get the data frame name without the "class." prefix
  df_name_without_prefix <- sub("^class\\.", "", df_name)
  
  # Iterate through each row in the data frame
  for (row in 1:nrow(current_df)) {
    filename <- current_df[row, "img_file_name"]
    
    # Construct the full path to the image file in the main directory
    full_path <- file.path(path_to_files, filename)
    
    # Construct the full path to the image file to be copied in the selected directory 
    selected_path <- file.path(selected_destination_dir_all, basename(filename))
    
    # Check if the file exists before copying
    if (file.exists(full_path)) {
      # Copy the image to the 'selected_per_class' directory
      file.copy(full_path, selected_path)
    } else {
      # Add the missing file to the list
      missing_files <- c(missing_files, filename)
    }
  }
}

# Print the list of missing files
if (length(missing_files) > 0) {
  cat("The following files were not found:\n")
  print(missing_files)
} else {
  cat("All files were found and copied successfully.\n")
}

# Set path to main directory
path_to_dest_files <- "~/Desktop/selected/"

# List all image files in all folders within the main directory
image_dest_files <- list.files(path = path_to_dest_files, pattern = "\\.(png|jpg|jpeg|tif|tiff)$", recursive = TRUE, full.names = TRUE)

# Count the number of image files
num_image_dest_files <- length(image_dest_files)

# Print the count of image files
cat("Number of image files in destination folder:", num_image_dest_files, "\n")
```

# Loop through each data frame (lists of annotated images) and copy them in dedicated folders
```{r}
# Initialize an empty list to keep track of missing files
missing_files <- list()

# Loop through each data frame
for (df_name in data_frame_names) {
  # Get the data frame using the name
  current_df <- get(df_name)
  
  # Rename filename column if needed
  current_df <- current_df %>% rename(img_file_name = V1)
  
  # Get the data frame name without the "class." prefix
  df_name_without_prefix <- sub("^class\\.", "", df_name)
  
  # Create directory for the current taxa list under the base selected directory
  selected_destination_dir <- file.path(selected_base_dir, df_name_without_prefix)
  if (!dir.exists(selected_destination_dir)) {
    dir.create(selected_destination_dir)
  }
  
  # Iterate through each row in the data frame
  for (row in 1:nrow(current_df)) {
    filename <- current_df[row, "img_file_name"]
    
    # Construct the full path to the image file in the main directory
    full_path <- file.path(path_to_files, filename)
    
    # Construct the full path to the image file in the selected directory 
    selected_path <- file.path(selected_destination_dir, basename(filename))
    
    # Check if the file exists before copying
    if (file.exists(full_path)) {
      # Copy the image to the 'selected_per_class' directory
      file.copy(full_path, selected_path)
    } else {
      # Add the missing file to the list
      missing_files <- c(missing_files, filename)
    }
  }
}

# Print the list of missing files
if (length(missing_files) > 0) {
  cat("The following files were not found:\n")
  print(missing_files)
} else {
  cat("All files were found and copied successfully.\n")
}

# Set path to main directory
path_to_sel_files <- "~/Desktop/selected_per_class/"

# List all image files in all folders within the main directory
image_sel_files <- list.files(path = path_to_sel_files, pattern = "\\.(png|jpg|jpeg|tif|tiff)$", recursive = TRUE, full.names = TRUE)

# Count the number of image files
num_image_sel_files <- length(image_sel_files)

# Print the count of image files
cat("Number of image files in destination folder:", num_image_sel_files, "\n")
```


