---
title: "Image Distributor"
output: null
---

This script copies annotated images from the main data folder (cpics_img) into corresponding folders in another directory: diatoms, acantharea, copepods, etc. 
The code also allows to copy all annotated images into a single folder (not aggregated by class).

Created: June 18th 2024
By: Enrique Montes

# Load taxa lists and metadata table
```{r}
library(tidyverse)
library(lubridate)

# specify the directory where the files are located
# dir_path <- "~/Library/CloudStorage/GoogleDrive-enriquemontes01@gmail.com/My Drive/GDrive/OCED_AOML/WS_cruises/plankton_imaging/CPICS/TS.Master_selection"
dir_path <- "/Users/enrique.montes/Desktop/cpics_img/TS.Master"

# obtain a list of file names in the directory
file_names <- list.files(path = dir_path, pattern = ".txt", full.names = TRUE)

# loop over each file and import the tables (use this for DATES)
for (file in file_names) {
  table_name <- gsub(".txt", "", basename(file)) # get the name of the table from the file name
  assign(table_name, read.table(file = file, header = FALSE, sep = "\t") %>%
           mutate(date = as.POSIXct(substr(V1, start = 24, stop = 38), format="%Y%m%d_%H%M%S", tz="UTC")))
}

# List of folder names with files lists
data_frame_names <- ls(pattern = "^class\\.")  # Get all data frames starting with "class."
```

# Set path to and create main directories (base directory where all images are and destination directory)
```{r}
path_to_files <- "~/Desktop/cpics_img/" # Base directory for all images
# selected_base_dir <- "~/Desktop/selected_per_class/"  # Base directory for selected files in separate folders
selected_destination_dir_all <- "~/Desktop/selected/"  # Base directory for selected files in common folder

# Use this path for creating a training dataset for the UW classifies
selected_base_dir <- "~/Desktop/uw_classifier/training_dataset/"  # Base directory for selected files in separate folders

# Create the base directory for selected files if it doesn't exist
if (!dir.exists(selected_base_dir)) {
  dir.create(selected_base_dir)
}

if (!dir.exists(selected_destination_dir_all)) {
  dir.create(selected_destination_dir_all)
}

# List all PNG image files in all folders within the main directory
image_files <- list.files(path = path_to_files, pattern = "\\.(png)$", recursive = TRUE, full.names = TRUE)

# Count the number of image files
num_image_files <- length(image_files)

# Print the count of image files
cat("Number of image files in base directory:", num_image_files, "\n")
```

# Loop through each data frame (lists of annotated images) and copy them all in a common folder
```{r}
# Initialize an empty list to keep track of missing files
missing_files <- list()

# Loop through each data frame
for (df_name in data_frame_names) {
  # Get the data frame using the name
  current_df <- get(df_name)
  
  # Rename filename column if needed
  current_df <- current_df %>% rename(img_file_name = V1)
  
  # Get the data frame name without the "class." prefix
  df_name_without_prefix <- sub("^class\\.", "", df_name)
  
  # Filter out rows with 'date' values prior to 2025-01-01
  current_df <- current_df %>% filter(date >= as.POSIXct("2024-11-01", tz="UTC") & date <= as.POSIXct("2024-11-30", tz="UTC"))
  
  # Skip the loop if no rows are left after filtering
  if (nrow(current_df) == 0) {
    next
  }

  # Iterate through each row in the data frame
  for (row in 1:nrow(current_df)) {
    filename <- current_df[row, "img_file_name"]
    
    # Skip if filename is NA or empty
    if (is.na(filename) || filename == "") {
      next
    }
    
    # Construct the full path to the image file in the main directory
    full_path <- file.path(path_to_files, filename)
    
    # Construct the full path to the image file to be copied in the selected directory 
    selected_path <- file.path(selected_destination_dir_all, basename(filename))
    
    # Check if the file exists before copying
    if (file.exists(full_path)) {
      # Copy the image to the 'selected_per_class' directory
      file.copy(full_path, selected_path)
    } else {
      # Add the missing file to the list
      missing_files <- c(missing_files, filename)
    }
  }
}

# Print the list of missing files
if (length(missing_files) > 0) {
  cat("The following files were not found:\n")
  print(missing_files)
} else {
  cat("All files were found and copied successfully.\n")
}

# Set path to main directory
  # path_to_dest_files <- "~/Desktop/selected/"

  # Use this path for creating a training dataset for the UW classifier
  path_to_dest_files <- "~/Desktop/uw_classifier/training_dataset_all/"

# List all image files in all folders within the main directory
image_dest_files <- list.files(path = path_to_dest_files, pattern = "\\.(png|jpg|jpeg|tif|tiff)$", recursive = TRUE, full.names = TRUE)

# Count the number of image files
num_image_dest_files <- length(image_dest_files)

# Print the count of image files
cat("Number of image files in destination folder:", num_image_dest_files, "\n")
```

# Loop through each data frame (lists of annotated images) and copy them in dedicated folders
```{r}
# Initialize an empty list to keep track of missing files
missing_files <- list()

# Loop through each data frame
for (df_name in data_frame_names) {
  # Get the data frame using the name
  current_df <- get(df_name)
  
  # Rename filename column if needed
  current_df <- current_df %>% rename(img_file_name = V1)
  
  # Get the data frame name without the "class." prefix
  df_name_without_prefix <- sub("^class\\.", "", df_name)
  
  # Filter out rows with 'date' values outside selected time period (e.g. from 2024-11-01 to 2024-11-30)
  # current_df <- current_df %>% filter(date >= as.POSIXct("2024-11-01", tz="UTC") & date <= as.POSIXct("2024-11-30", tz="UTC"))
  
  # Skip the loop if no rows are left after filtering
  if (nrow(current_df) == 0) {
    next
  }

  # Create directory for the current taxa list under the base selected directory
  selected_destination_dir <- file.path(selected_base_dir, df_name_without_prefix)
  if (!dir.exists(selected_destination_dir)) {
    dir.create(selected_destination_dir)
  }
  
  # Iterate through each row in the data frame
  for (row in 1:nrow(current_df)) {
    filename <- current_df[row, "img_file_name"]
    
    # Skip if filename is NA or empty
    if (is.na(filename) || filename == "") {
      next
    }
    
    # Construct the full path to the image file in the main directory
    full_path <- file.path(path_to_files, filename)
    
    # Construct the full path to the image file in the selected directory 
    selected_path <- file.path(selected_destination_dir, basename(filename))
    
    # Check if the file exists before copying
    if (file.exists(full_path)) {
      # Copy the image to the 'selected_per_class' directory
      file.copy(full_path, selected_path)
    } else {
      # Add the missing file to the list
      missing_files <- c(missing_files, filename)
    }
  }
}

# Print the list of missing files
if (length(missing_files) > 0) {
  cat("The following files were not found:\n")
  print(missing_files)
} else {
  cat("All files were found and copied successfully.\n")
}

# Set path to main directory
  # path_to_sel_files <- "~/Desktop/selected_per_class/"
  
  # Use this path for creating a training dataset for the UW classifier
  path_to_sel_files <- "~/Desktop/uw_classifier/training_dataset/"

# List all image files in all folders within the main directory
image_sel_files <- list.files(path = path_to_sel_files, pattern = "\\.(png|jpg|jpeg|tif|tiff)$", recursive = TRUE, full.names = TRUE)

# Count the number of image files
num_image_sel_files <- length(image_sel_files)

# Print the count of image files
cat("Number of image files in destination folder:", num_image_sel_files, "\n")
```


# Check for repeated files in 'selected_per_class' and 'selected' folders. This can happen if the same image is annotated as two or more classes, say as nauplia and copepod.
```{r}
# Function to extract the last 25 characters from file names
get_last_25 <- function(file_names) {
  sapply(file_names, function(name) substr(name, pmax(1, nchar(name) - 24), nchar(name)))
}

# Apply the function to image_sel_files (selected_per_class) and image_dest_files (selected)
sel1_last_25 <- as.data.frame(get_last_25(image_sel_files)) 
colnames(sel1_last_25) <- "fname_sel1"
sel2_last_25 <- as.data.frame(get_last_25(image_dest_files))
colnames(sel2_last_25) <- "fname_sel2"

# Look for files in one directory not present in the other
unaccounted_for <- setdiff(sel2_last_25$fname_sel2, sel2_last_25$fname_sel2)
cat("Unaccounted for image files in selected folder:", unaccounted_for, "\n")

# Check for duplicate files
repeated_files <- sel1_last_25$fname_sel1[duplicated(sel1_last_25$fname_sel1)]
cat("Duplicate image files in directory:", repeated_files, "\n")

# List all folders in the directory
folders <- list.dirs(selected_base_dir, full.names = TRUE, recursive = FALSE)

# Initialize a list to store folders containing repeated files
folders_with_repeats <- list()

# Loop through each folder to check for repeated files
for (folder in folders) {
  # List all files in the folder
  files_in_folder <- list.files(folder, full.names = FALSE)
  
  # Extract the last 25 characters of each file name
  last_25_in_folder <- sapply(files_in_folder, function(name) {
    substr(name, pmax(1, nchar(name) - 24), nchar(name))
  })
  
  # Check if any of these file names match repeated_files
  repeats_in_folder <- intersect(last_25_in_folder, repeated_files)
  
  # If there are matches, add the folder to the list
  if (length(repeats_in_folder) > 0) {
    folders_with_repeats[[folder]] <- repeats_in_folder
  }
}

# Display folders with repeated files
cat("Folders containing repeated files:\n")
print(folders_with_repeats)

```

# # Aggregate all or selected unclassified PNG files in a single folder to classify using the UW classifier
```{r}
# R Script to Copy and Filter PNG Files by Datetime
#
# This script recursively finds all PNG files in 'cpics_img',
# extracts the datetime from each filename, filters them based on a
# specified date range, and copies the matching files to a
# destination folder.

library(dplyr)
library(fs)

# The main folder containing all your date-stamped subfolders
source_directory <- "~/Desktop/cpics_img/"

# The single folder where you want to copy the filtered PNG files
destination_directory <-"~/Desktop/uw_classifier/unclassified_dataset/unclassified/"

# Define the date range for filtering (inclusive)
# The script will keep files with dates ON or BETWEEN these two dates.
start_date <- as.POSIXct("2025-01-01 00:00:00", tz = "UTC")
end_date <- as.POSIXct("2025-02-01 00:00:00", tz = "UTC")

# Step 1: Find all PNG files recursively in the source directory
cat("Finding all PNG files in '", source_directory, "'...\n", sep = "")
all_files <- dir_ls(
  path = source_directory,
  recurse = TRUE,
  type = "file",
  regexp = "\\.png$" # Use a regular expression to match files ending in .png
)

if (length(all_files) == 0) {
  stop("No PNG files were found in the source directory. Please check the path.")
}

cat("Found", length(all_files), "total PNG files.\n")

# Step 2: Create a data frame with file paths and extracted datetimes
# The filename format is assumed to be 'YYYYMMDD_hhmmss.sss.n.png'
# We extract the 'YYYYMMDD_hhmmss' part to create a datetime object.
files_df <- tibble(
  file_path = all_files
) %>%
  mutate(
    # Extract the first 15 characters from the filename (e.g., "20250516_011837")
    datetime_str = substr(path_file(file_path), 1, 15),
    # Convert the string to a proper POSIXct datetime object
    date = as.POSIXct(datetime_str, format = "%Y%m%d_%H%M%S", tz = "UTC")
  )

cat("Successfully extracted datetimes from filenames.\n")

# Step 3: Filter the data frame to keep only files within the desired date range
cat("Filtering files between", format(start_date, "%Y-%m-%d"), "and", format(end_date, "%Y-%m-%d"), "...\n")

filtered_files_df <- files_df %>%
  filter(
    date >= start_date & date <= end_date
  )

if (nrow(filtered_files_df) == 0) {
  cat("No files matched the specified date range. No files will be copied.\n")
} else {
  cat("Found", nrow(filtered_files_df), "files within the date range.\n")
  
  # Step 4: Copy the filtered files to the destination directory
  cat("Copying files to '", destination_directory, "'...\n", sep = "")
  
  # The `file_copy` function from 'fs' can handle vectors of files easily.
  # It copies the files from the 'file_path' column into the destination folder.
  file_copy(
    path = filtered_files_df$file_path,
    new_path = destination_directory,
    overwrite = TRUE # Set to FALSE if you don't want to overwrite existing files
  )
  
  cat("\n--- All Done! ---\n")
  cat(nrow(filtered_files_df), "files have been successfully copied to:\n")
  cat(path_real(destination_directory), "\n")
}

# # NOTE: use this command in terminal to upload 'unclassified_dataset' to Holocron: rsync -avz /Users/enrique.montes/Desktop/uw_classifier/unclassified_dataset enrique.montes@holocron.aoml.noaa.gov:/home/enrique.montes@CNS.local/plankton_imaging


```

