---
title: "seascape_similarity_coefficients"
output: none
date: "2025-07-10"
---

# # Calculate Morisita-Horn similarity (C-MH; Chao et al 2006)  and other indices for each seascape using plankton groups

# Create data frame: Run the first two chunks '1 and 2 of occurrence_counts.Rmd' to create 'taxa_meta'
```{r}
# Create data frame using count data (not concentration!)

rm(list = ls()[ls() != "taxa_meta"])

c_mh <- taxa_meta %>%
  # mutate(across(c(Centric,Ceratium,Chaetoceros,Chain2,Chain3,Guinardia,Neocalyptrella,
  #                 Noctiluca,Tricho))) %>%
  mutate(across(c(Acantharea,Chaetognaths,Ostracods,Copepods,Decapods,Echinoderms,Jellies,
    Larvaceans,Polychaets,Pteropods))) %>%
  select(Station, 
         year, 
         month,  
         X8.day.seascapes,
         focus_threshold,
         sampled_vol_per_min,
         total_vol_sampled,
         Centric,
         Ceratium,
         Chaetoceros,
         Chain2,
         Chain3,
         Guinardia,
         Neocalyptrella,
         Noctiluca,
         Tricho) %>%
         # Acantharea,
         # Chaetognaths,
         # Ostracods,
         # Copepods,
         # Decapods,
         # Echinoderms,
         # Jellies,
         # Larvaceans,
         # Polychaets,
         # Pteropods) %>%
  filter(!is.na(total_vol_sampled)) %>%
  filter(!is.na(X8.day.seascapes))

# # Transform taxa_meta_concentration to long format
c_mh_long <- c_mh %>%
pivot_longer(cols = c(Centric,Ceratium,Chaetoceros,Chain2,Chain3,Guinardia,Neocalyptrella,
                  Noctiluca,Tricho),
  # pivot_longer(cols = c(Acantharea, Chaetognaths,Ostracods,Copepods,Decapods,
  #                       Echinoderms, Jellies, Larvaceans,Polychaets, Pteropods),
               names_to = "species", values_to = "species_concentration")

```

# # Step 1: Create a meta-community for each seascape: aggregate all data for each seascape category into a single, representative community.
```{r}
seascape_meta_communities <- c_mh_long %>%
  # Group by seascape and species to pool all data
  group_by(X8.day.seascapes, species) %>%
  # Sum all concentrations to get a total abundance for each species
  summarise(abundance = sum(`species_concentration`, na.rm = TRUE), .groups = 'drop') %>%
  # Remove any records with zero abundance
  filter(abundance > 0)

```

# # Step 2: Prepare for pairwise comparisons: nest the data so each seascape's community can be treated as a single object, then create all unique pairs of seascapes to compare.
```{r}
# Nest the community data
nested_seascapes <- seascape_meta_communities %>%
  group_by(X8.day.seascapes) %>%
  nest(community_data = -X8.day.seascapes)

# Create a dataframe with all pairs of seascapes
seascape_pairs <- nested_seascapes %>%
  select(seascape_1 = X8.day.seascapes, data_1 = community_data) %>%
  crossing(
    nested_seascapes %>%
      select(seascape_2 = X8.day.seascapes, data_2 = community_data)
  )
```

# # Step 3: Calculate C_MH for all pairs using the same robust calculation function applied it to every pair.
```{r}
# The Morisita-Horn calculation function
calculate_morisita_horn <- function(df1, df2) {
  combined <- full_join(df1, df2, by = "species", suffix = c("_1", "_2")) %>%
    mutate(
      abundance_1 = replace_na(abundance_1, 0),
      abundance_2 = replace_na(abundance_2, 0)
    )
  n <- sum(combined$abundance_1)
  m <- sum(combined$abundance_2)
  if (n == 0 || m == 0) return(NA_real_)
  p_i <- combined$abundance_1 / n
  q_i <- combined$abundance_2 / m
  numerator <- 2 * sum(p_i * q_i)
  denominator <- sum(p_i^2) + sum(q_i^2)
  return(numerator / denominator)
}

# Apply the function to every pair
pairwise_similarity <- seascape_pairs %>%
  rowwise() %>%
  mutate(C_MH = calculate_morisita_horn(data_1, data_2)) %>%
  select(seascape_1, seascape_2, C_MH)
```

# # Step 4: Create a C_MH Heatmap
```{r}
ggplot(pairwise_similarity, aes(x = factor(seascape_1), y = factor(seascape_2), fill = C_MH)) +
  geom_tile(color = "white") +
  # Add the similarity values as text on the tiles
  geom_text(aes(label = round(C_MH, 2)), color = "black", size = 4) +
  # Use a color scale that shows high similarity as "hot"
  scale_fill_gradient(low = "lightblue", high = "darkblue", name = "C_MH\nSimilarity") +
  labs(
    title = "Pairwise Morisita-Horn Similarity Between Seascapes",
    x = "Seascape",
    y = "Seascape"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


# # Step 5: The Chao-Jaccard calculation function: calculate the abundance-based Jaccard similarity index, which gives stronger weight to rare species.
```{r}
# # # Use this for Chao-Jaccard abundance based index
calculate_chao_jaccard <- function(df1, df2) {
  # Join data to get a complete species list
  combined <- full_join(df1, df2, by = "species", suffix = c("_1", "_2")) %>%
    mutate(
      abundance_1 = replace_na(abundance_1, 0),
      abundance_2 = replace_na(abundance_2, 0)
    )

  # Calculate total abundances
  n <- sum(combined$abundance_1)
  m <- sum(combined$abundance_2)
  if (n == 0 || m == 0) return(NA_real_)

  # Calculate relative abundances
  p_i <- combined$abundance_1 / n
  q_i <- combined$abundance_2 / m

  # Calculate the sum of minimums and maximums
  sum_of_mins <- sum(pmin(p_i, q_i))
  sum_of_maxs <- sum(pmax(p_i, q_i))

  # The Chao-Jaccard (abundance-based) similarity is sum(min) / sum(max)
  similarity <- sum_of_mins / sum_of_maxs

  return(similarity)
}

################################################################################
################################################################################

# # # Use this for Jaccard abundance based index
# calculate_abundance_jaccard <- function(df1, df2) {
#   # 1. Calculate the total abundance (or concentration) for each community
#   n <- sum(df1$abundance)
#   m <- sum(df2$abundance)
#   
#   # Return NA if either community is empty
#   if (n == 0 || m == 0) return(NA_real_)
# 
#   # 2. Identify only the species shared between both communities
#   shared_species <- inner_join(df1, df2, by = "species", suffix = c("_1", "_2"))
# 
#   # 3. Calculate U and V as defined
#   # U = Sum of relative abundances of shared species in community 1
#   U <- sum(shared_species$abundance_1) / n
#   
#   # V = Sum of relative abundances of shared species in community 2
#   V <- sum(shared_species$abundance_2) / m
# 
#   # 4. Apply the final Jaccard equation
#   numerator <- U * V
#   denominator <- U + V - (U * V)
#   
#   # Handle the case of no shared species (U and V are 0)
#   if (denominator == 0) {
#     return(0)
#   }
#   
#   similarity <- numerator / denominator
#   return(similarity)
# }

```

# # Step 6: Calculate pairwise similarity: create all seascape pairs and apply the new function.
```{r}
# Nest the community data
nested_seascapes <- seascape_meta_communities %>%
  group_by(X8.day.seascapes) %>%
  nest(community_data = -X8.day.seascapes)

# Create a dataframe with all pairs of seascapes
seascape_pairs <- nested_seascapes %>%
  select(seascape_1 = X8.day.seascapes, data_1 = community_data) %>%
  crossing(
    nested_seascapes %>%
      select(seascape_2 = X8.day.seascapes, data_2 = community_data)
  )

# Apply the function to every pair
pairwise_similarity <- seascape_pairs %>%
  rowwise() %>%
  mutate(C_J = calculate_chao_jaccard(data_1, data_2)) %>%
  # mutate(C_J = calculate_abundance_jaccard(data_1, data_2)) %>%
  select(seascape_1, seascape_2, C_J)
```

# # Step 7: Create the Chao-Jaccard Heatmap
```{r}
ggplot(pairwise_similarity, aes(x = factor(seascape_1), y = factor(seascape_2), fill = C_J)) +
  geom_tile(color = "white") +
  geom_text(aes(label = round(C_J, 2)), color = "black", size = 4) +
  scale_fill_gradient(low = "#fee8c8", high = "#b30000", name = "Chao-Jaccard\nSimilarity") +
  labs(
    title = "Pairwise Chao-Jaccard Similarity Between Seascapes",
    x = "Seascape",
    y = "Seascape"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Replace C_J with NA where seascape_2 < seascape_1
pairwise_similarity_updated <- pairwise_similarity %>%
  mutate(
    C_J = if_else(
      # Condition: is the numeric value of seascape_2 less than seascape_1?
      as.numeric(as.character(seascape_2)) < as.numeric(as.character(seascape_1)),
      # If TRUE, replace with NA
      NA_real_,
      # If FALSE, keep the original C_J value
      C_J
    )
  )

pairwise_similarity_updated$seascape_1 <- factor(pairwise_similarity_updated$seascape_1,
                                   levels = c(3, 11, 13, 15, 21, 27),
                                   labels = c("TST", "TSU", "SGMI", "TS", "WBHN", "HE"))
pairwise_similarity_updated$seascape_2 <- factor(pairwise_similarity_updated$seascape_2,
                                   levels = c(3, 11, 13, 15, 21, 27),
                                   labels = c("TST", "TSU", "SGMI", "TS", "WBHN", "HE"))
# View the updated data frame
similarity_hm <- ggplot(pairwise_similarity_updated, aes(x = factor(seascape_1), y = factor(seascape_2), fill = C_J)) +
  geom_tile(color = "white") +
  geom_text(aes(label = round(C_J, 2)), color = "black", size = 8) +
  scale_fill_gradient(
    low = "#fee8c8", 
    high = "#b30000", # b30000 for zooplankton / 006400 for phyto
    name = "Similarity", 
    limit = c(0.21,1),
    breaks = seq(0.2, 1.0, by = 0.2)) +  
  labs(x = NULL, y = NULL) +
  theme_minimal() +
  theme(
    axis.text.x  = element_text(size = 32, angle = 0, color = "black"),
    axis.text.y  = element_text(size = 32, angle = 0, color = "black"),
    panel.grid = element_blank(),
    legend.text = element_text(size = 18),
    legend.title = element_text(size = 22, margin = margin(b = 15)))
similarity_hm

# ggsave("seascape_similarity_zoopl.svg", plot = similarity_hm, width = 16, height = 8, device = "svg")
```

